---
title: "Assignment 3"
author: "Nathaniel Fleming"
date: "2023-04-03"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
Data Background:

This data contains information on on each pokemon from the first generation Kanto Region to the 6th generation Kalos Region Pokemon X and Y.  The data contains 721 unique pokemon with 800 entries for pokemon. Some of these entries include the same pokemon but with a region/generation specific entry, these values will not be included for complexities sake. 

The fields include:

X: ID for each pokemon

Name: Name of each pokemon

Type 1: Each pokemon has a type, this determines weakness/resistance to attacks

Type 2: Some pokemon are dual type and have 2

Total: sum of all stats that come after this, a general guide to how strong a pokemon is

HP: hit points, or health, defines how much damage a pokemon can withstand before fainting

Attack: the base modifier for normal attacks (eg. Scratch, Punch)

Defense: the base damage resistance against normal attacks

SP Atk: special attack, the base modifier for special attacks (e.g. fire blast, bubble beam)

SP Def: the base damage resistance against special attacks

Speed: determines which pokemon attacks first each round.

Legendary: These pokemon can only be caught once and usually have higher stats 

Through this dataset I want to observe what features have an impact on the total stats a pokemon has. through this I will be conducting analysis on several different stats and types. 


Import the libraries for the analysis 

```{r,echo=FALSE}
library(ggplot2)
file_path <- "C:/Users/Nathaniel Fleming/Desktop/Mohawk/Coding/Assignment 3/"

```

Import the dataset, I chose pokemon to be the dataset to analyze as it is near and dear to my heart. The first step is to read the data.
```{r}
pokemonData<- read.csv("C:/Users/Nathaniel Fleming/Desktop/Mohawk/Coding/Assignment 3/Pokemon/Pokemon.csv")
```
I imported the correct colour pallete for the pokemon types so that visualization is easier and more accurate to the colour of the types that are in pokemon. 
```{r}
# Define custom color palette
pokemon_colors <- c("Fire" = "#EE8130", "Water" = "#6390F0", "Grass" = "#7AC74C", "Electric" = "#F7D02C", 
                    "Ice" = "#96D9D6", "Fighting" = "#C22E28", "Poison" = "#A33EA1", "Ground" = "#E2BF65", 
                    "Flying" = "#A98FF3", "Psychic" = "#F95587", "Bug" = "#A6B91A", "Rock" = "#B6A136", 
                    "Ghost" = "#735797", "Dragon" = "#6F35FC", "Dark" = "#705746", "Steel" = "#B7B7CE", 
                    "Fairy" = "#D685AD")

```
Explore the dataset and get to know the variables and datatypes associated with the csv
```{r}
View(pokemonData)
```
Checking the structure of the data 
```{r}
str(pokemonData)
```
Checking the summary statistics of the dataset
```{r}
summary(pokemonData)
```
```{r}
head(pokemonData)
tail(pokemonData)
```

Cleaning the Dataset
```{r}
dupe_rows<- duplicated(pokemonData$X.)
pokemonNoDupes<- subset(pokemonData,!dupe_rows)
summary(pokemonNoDupes)
```
Check for outliers
```{r}

# Create a box plot of Total stats by Type
boxplot_type_total <- ggplot(pokemonNoDupes, aes(x = Type.1, y = Total, fill = Type.1)) +
  geom_boxplot() +
  labs(x = "Type", y = "Total Stats", title = "Total Stats by Type")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Apply custom color palette
boxplot_type_total + scale_fill_manual(values = pokemon_colors)
ggsave(file.path(file_path, "boxplot.png"), plot = boxplot_type_total)

```
Normal type pokemon seem to have one of the larger variance in stat types than 

I wanted to compare pokemon that have more than one type to those with only one so I subset the data into multitype and single type pokemon. This way we can compare pokemon that have more than one type versus the pokemon with one type. 

```{r}
# Filter out blank values that are just spaces in the Type.2 column
multi_type_pokemon <- pokemonNoDupes[nchar(trimws(pokemonNoDupes$Type.2)) > 0, ]

# Filter out blank values in the Type.2 column
multi_type_pokemon

#Single Type pokemon 
# Filter for empty values in the Type.2 column
Single_Type_Pokemon <- pokemonNoDupes[is.na(pokemonNoDupes$Type.2) | nchar(trimws(pokemonNoDupes$Type.2)) == 0, ]
Single_Type_Pokemon

```



```{r}
# Create scatterplot with custom colors
scatterplot<-ggplot(pokemonNoDupes, aes(x = Attack, y = Defense, color = Type.1)) +
  geom_point() +
  scale_color_manual(values = pokemon_colors)
ggsave(file.path(file_path, "scatterplot.png"), plot = scatterplot)

```
Exploration into the speeds 
```{r}
speed_hist <- ggplot(pokemonNoDupes, aes(x = Speed)) +
  geom_histogram(bins = 20) +
  labs(x = "Pokemon Attack Speed", y = "Count", title = "Distribution of Pokemon Attack Speeds")
speed_hist
# Save the plot to a PNG file
ggsave(file.path(file_path, "speed_hist.png"), plot = speed_hist)
```

```{r}
legendary_pokemon <- subset(pokemonNoDupes,Legendary == 'True' )
legendary_pokemon

```

```{r}
Legendary_atk_vs_def <- ggplot(legendary_pokemon, aes(x = Attack, y = Defense, color = Type.1)) + 
  geom_point() +
  labs(title = "Attack vs Defense for Legendary Pokemon", 
       x = "Attack", y = "Defense") +
   scale_color_manual(values = pokemon_colors)
ggsave(file.path(file_path,"legendary_atk_vs_def.png"),plot = Legendary_atk_vs_def)

```

```{r}


# calculate frequency of each type
type_freq <- table(pokemonNoDupes$Type.1)

# convert table to data frame and sort by frequency
type_freq_df <- data.frame(Type = names(type_freq), Freq = as.vector(type_freq))
type_freq_df <- type_freq_df[order(type_freq_df$Freq, decreasing = TRUE), ]

# create bar plot
Type_Freq <- ggplot(type_freq_df, aes(x = reorder(Type, -Freq), y = Freq, fill = Type)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = pokemon_colors) +
  xlab("Type") +
  ylab("Frequency") +
  ggtitle("Frequency of Pokemon Types") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5),legend.position = 'none')
  
  ggsave(file.path(file_path,"Pokemon_Freq_by_Type.png"),plot = Type_Freq)

```

